---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# HeartR package

## Description

__HeartR__ is an R package designed for educational purposes to explore, summarise and simulate heart disease related data. Informative plots may be created to further analyse the distribution of variables and correlation between features. The package provides several functions and one S3 method: 

  1. `plot_heart`: a function that provides multiple visualisation options
  2. `simulation`: for generating a synthetic heart disease dataset. However it may be used for other datasets with binary outcomes
  3. `summ_heartr`: for creating a list of summary statistics that will be used further in the `print.summ_heartr` method
  4. `plot_summary`: for generating interactive plots of summary statistics
  5. An S3 `print.summ_heartr` method: for providing an appealing customisable table for the output of the summ_heartr function

## Installation

You can install the development version of HeartR from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
# install.packages("pak")
pak::pak("faridat45/HeartR")
```

## Example

This is a basic example which shows you how to generate synthetic data, adjusting the categorical and numeric variables included and number of observations:

```{r example}
library(HeartR)

sim_data <- simulation(
  data = heart_dat,
  n = 20,
  contVars = c("age", "cholesterol","restingBP"),
  catVars = c("sex","chestPT","exerAngina")
)
utils::head(sim_data)
```

From this dataset, plots can be generated using the `plot_heart` function:

```{r pressure, echo = FALSE, warning=FALSE}
plot_heart(sim_data, "boxplot", var_x = "age", var_y = "target")
```

A more detailed introduction is provided in the heartR vignette
